### 2023-07-20 chapters/background.tex
Add benchmarking paper
Updates
Add WIP for introduction and background

\chapter[background]{Background}\label{chp:background}

\begin{enumerate}
    \item Variational inference and variational autoencoders
    \item Automatic speech recognition
    \item Self-supervised learning
    \item 
\end{enumerate}

@@

\cite{bahdanau_neural_2014}

\begin{enumerate}
    \item Introduce variational inference in relation to variational autoencoders
    \item Introduce ASR and CTC
\end{enumerate}

>>>>

%!TEX root = ../thesis.tex

% ~8 pages

\chapter[technical background]{Technical Background}\label{chp:technical-background}

\section{Uncertainty}

% https://en.wikipedia.org/wiki/Uncertainty#cite_note-3
% https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=8371683

In the general public, the concept of uncertainty goes by many names and its meaning can be dependent on a specific context. However, in most quantitative scientific fields, there seems to be a consensus definition along the lines of the following \cite{hubbard_how_2014}:

\begin{center}
    \textit{Uncertainty is the lack of certainty; a state of limited knowledge where it is impossible to exactly describe the existing state, a future outcome, or more than one possible outcome.}
\end{center}

While a definition of uncertainty made using only words easily begs philosophical inquisition, information theory provides a rigourous method for quantifying uncertainty \cite{kabir_neural_2018} by the language of probabilities. 
In this context, uncertainty is understood as the information entropy of a random variable \cite{mackay_information_2003}.

\subsection{Information entropy}
First characterised by Claude Shannon in 1948 \cite{shannon_mathematical_1948}, information entropy $H(X)$ is a measure of the average amount of information contained in a random variable $X$. Shannon's definition follows from three fundamental axioms of information theory. Let $I_X(x)$ be the information carried by a specific outcome $x$ when sampling the random variable $X$ with probability mass function $p_X(x)$. Then, the axioms are as follows:
\begin{enumerate}
    \item  $I_X(x)$ is a monotonically decreasing function in $p_X(x)$. The more likely an outcome is, the less information it carries.
    \item Outcomes that are certain to happen carry no information. $I_X(1) = 0$.
    \item If an outcome $x_i$ is independent of another outcome $x_j$, then $I(x_i x_j) = I(x_i) + I(x_j)$. The information carried by independent events is the sum of the information carried by each event.
\end{enumerate}

From these axioms, Shannon derived the following expression for the information content of an event $x$ with probability $p_X(x)$:
\begin{equation}
    I_X(x) = -\log_b p_X(x)
\end{equation}

information entropy of a random variable $X$ with $n$ possible outcomes:
\begin{equation}
    H(X) = -\sum_{i=1}^n p_i \log p_i
\end{equation}

% Two kinds of uncertainty:\\
% - Aleatoric: Uncertainty inherent to the data. Irreducible. "Known unknowns".\\
% - Epistemic: Uncertainty due to lack of knowledge. Can be reduced with more data. "Unknown unknowns".\\

% https://arxiv.org/pdf/2306.05674.pdf

% We compare our framework with several major related lines of work. First, our work focuses on the quantification of epistemic uncertainty, which refers to the errors coming from the inadequacy of the model or data noises. This is different from aleatoric uncertainty, which refers to the intrinsic stochasticity of the problem [87, 91, 112, 59, 34], or predictive uncertainty which captures the sum of epistemic and aleatoric uncertainties (but not their dissection) [89, 92, 12, 3, 22]. Regarding epistemic uncertainty, a related line of study is deep ensemble that aggregates predictions from multiple independent training replications [76, 70, 38, 8, 89]. This approach, as we will make clear later, can reduce and potentially quantify procedural variability, but a naive use would require demanding retraining effort and does not address data variability. Another line is the Bayesian UQ approach on neural networks [41, 2]. This regards network weights as parameters subject to common priors such as Gaussian. Because of the computation difficulties in exact inference, an array of studies investigate efficient approximate inference approaches to estimate the posteriors [40, 46, 16, 31, 30, 90, 74, 53]. While powerful, these approaches nonetheless possess inference error that could be hard to quantify, and ultimately finding rigorous guarantees on the performance of these approximate posteriors remains open to our best knowledge.

% 87: 
% 91: 
% 112:
% 59: 
% 34:

% 89: 
% 92: 
% 12:
% 3: 
% 22:

% 76: 
% 70: 
% 38: 
% 8: 
% 89:

\section{Probabilistic latent variable models}

\section{Speech representation learning}

Modelling paradigms
\begin{enumerate}
    \item Variational inference and variational autoencoders
    \item Automatic speech recognition
    \item Self-supervised learning
\end{enumerate}

@@

Two approaches to statistics:\\
- Bayesian\\
- Frequentist\\

Two kinds of uncertainty:\\
- Aleatoric: Uncertainty inherent to the data. Irreducible. "Known unknowns".\\
- Epistemic: Uncertainty due to lack of knowledge. Can be reduced with more data. "Unknown unknowns".\\

Approaches to uncertainty in deep learning:\\
- Tuning of supervised classifiers to output well-calibrated probabilities that correspond to the actual likelihood of the prediction being correct (on some validation set).\\
- Learning rich unsupervised representations of the data that can be used to estimate the uncertainty of the prediction (e.g. by measuring the distance to the nearest training example in the latent space).\\
- Learning of a probability distribution over the parameters of the model, which can then be used to compute the uncertainty of the prediction (Bayesian neural networks).\\

In the field of machine learning, uncertainty is often represented by a probability distribution. The most common approach is to use Bayesian methods, where uncertainty is captured by posterior distributions over model parameters and predictions. Bayesian neural networks, for instance, offer a powerful framework to model uncertainty in deep learning architectures. By incorporating prior beliefs and updating them based on observed data, these networks can produce probabilistic predictions that provide a measure of uncertainty. This is particularly useful in applications where high-confidence predictions are required, and the consequences of errors can be significant.

Recently, another promising approach to uncertainty estimation in machine learning is the use of Monte Carlo Dropout \cite{gal_dropout_2016}. Monte Carlo Dropout leverages the idea of dropout regularization, originally employed during training to prevent overfitting. To form a prediction with associated uncertainty, Monte Carlo Dropout proposes to make multiple forward passes through the network while sampling different dropout masks for each one. This leads to obtaining a distribution of outputs for each input sample and the variance among these sampled predictions serves as a measure of uncertainty. Monte Carlo Dropout has shown remarkable success in various tasks such as image classification, object detection, and natural language processing. It is computationally efficient and can be easily incorporated into existing deep learning architectures, making it an attractive choice for uncertainty estimation.

Furthermore, there has been a surge of interest in ensemble methods for uncertainty estimation. Ensembles combine the predictions of multiple models to obtain a more robust and calibrated uncertainty measure. Bagging and boosting techniques, which have been widely used in the field of statistics, have found their way into the realm of machine learning for uncertainty estimation as well. By training multiple models with different initializations or using diverse learning algorithms, ensembles can capture different sources of uncertainty, thereby providing a comprehensive assessment of the overall uncertainty in the predictions.

Despite the progress made in uncertainty estimation for machine learning models, challenges still remain. One key concern is the interpretability of uncertainty measures. While probabilistic outputs are more informative than point estimates, effectively communicating uncertainty to end-users and decision-makers is not trivial. Developing visualization techniques and intuitive explanations for uncertainty is an ongoing area of research. Additionally, quantifying uncertainty in high-dimensional data or complex models with massive amounts of parameters requires careful consideration and efficient computational strategies.

### 2023-07-20 chapters/introduction.tex
Add benchmarking paper
Updates
Add WIP for introduction and background

@@

>>>>

%!TEX root = ../thesis.tex

@@

% ~9 pages

Uncertainty is a fundamental part of human experience. Understanding and quantifying uncertainty is crucial in making informed decisions, drawing meaningful conclusions, and assessing the reliability of predictions made from observations. Yet, capturing it in a mathematical model is a difficult task.

Historically, uncertainty estimation has been a central topic in the field of statistics. Uncertainty has traditionally been addressed through methods like confidence intervals, hypothesis testing, and Bayesian inference. Confidence intervals provide a range of plausible values for a parameter, while hypothesis testing evaluates the significance of differences between groups or variables \cite{blitzstein_introduction_2019}. Bayesian inference, on the other hand, offers a powerful framework to incorporate prior knowledge and update beliefs based on observed data using probability theory. These approaches have been extensively studied and applied in various domains, from social sciences to natural sciences, to improve the robustness and generalizability of statistical analyses \cite{gelman_bayesian_2013}. However, the widespread adoption in diverse applications of machine learning, and in particular deep neural networks, has posed new challenges in dealing with uncertainty.

Machine learning algorithms are designed to learn patterns from data and make predictions or decisions based on those patterns. Since many applications of machine learning are concerned with real-world phenomena, it is important to understand the uncertainty associated with the predictions made by the model. For instance, in medical diagnosis, knowing the uncertainty in a model's prediction is crucial when the consequences of a false positive or false negative can be significant. In speech recognition, knowing that a certain word was hard to understand for the model can help avoid misinterpretation of the transcribed text.

Speech processing is

% The performance of machine learning models is highly dependent on the quality and quantity of data used for training. In many cases, the data used to train machine learning models are collected from the real world, and therefore, may contain biases and errors which models may learn to mirror. In addition, the complexity of modern machine learning models makes it difficult to understand how they arrive at their predictions. This is especially true for deep neural networks, which are often referred to as black-box models. The lack of interpretability makes it difficult to assess the reliability of the model's predictions.

% Modern speech processing relies on high-performance parallel computing [40, 157], large volumes of data [136, 213], and years of innovation in model design [109, 266].

% 40: High performance convolutional neural networks for document processing \cite{chellapilla_high_2006}
% 157: Imagenet classification with deep convolutional neural networks \cite{krizhevsky_imagenet_2012}
% 136: Libri-light: A benchmark for asr with limited or no supervision \cite{kahn_libri-light_2020}
% 213: Librispeech: an ASR corpus based on public domain audio books \cite{panayotov_librispeech_2015}
% 109:  Long short-term memory \cite{hochreiter_long_1997}
% 266: Attention Is All You Need \cite{vaswani_attention_2017}

The research in this thesis covers a number of different modelling approaches and tasks, including

The research in this thesis covers a varied set of topics. One group of the produced studies are concerned with generative latent variable models and their applications to uncertainty estimation and speech modelling \cite{havtorn_hierarchical_2021,havtorn_benchmarking_2022,bergamin_modelagnostic_2022}. Another group of studies are concerned with self-supervised speech representation learning and automatic speech recognition \cite{borgholt_scaling_2021,borgholt_we_2021,mohamed_selfsupervised_2022,borgholt_brief_2022}. A final group

This first part of the thesis ties together these different studies by providing a high-level overview of the research topic and the main contributions of the thesis.

covering both representation learning and uncertainty estimation and their applications to speech recognition and automated medical coding.

\subsection{Motivation: Speech recognition}

\subsection{Motivation: Automated medical coding}

\subsection{Challenges in uncertainty estimation}

General challenges in uncertainty estimation:
\begin{enumerate}
    \item \textbf{Accuracy}: The uncertainty estimation method should be accurate.
    \item \textbf{Interpretability}: The uncertainty estimation method should be interpretable.
    \item \textbf{Robustness}: The uncertainty estimation method should be robust to adversarial attacks.
    \item \textbf{Applicability}: The uncertainty estimation method should be applicable to a wide range of models without requiring extensive modifications.
    \item \textbf{Efficiency}: The uncertainty estimation method should be computationally efficient at training and prediction time.
    \item \textbf{Scalability}: The uncertainty estimation method should scale to large datasets.
\end{enumerate}

\begin{enumerate}
    \item How can we learn representations that can form the basis of both prediction and uncertainty estimation?
    \item 
\end{enumerate}

\subsection{Scope of the thesis}

\iffalse

Uncertainty is a fundamental part of human experience. Yet, capturing it in a mathematical model is a difficult task. Historically, uncertainty estimation has been a central topic in the field of statistics. While classical statistical methods have provided valuable insights, they often assume well-defined probability distributions and may not fully capture the complexities of uncertainty in real-world scenarios. Moreover, they tend to rely heavily on parametric assumptions, which might not hold true for complex data with high-dimensional features and non-linear relationships. Therefore, in recent years, researchers have turned their attention towards more flexible and data-driven approaches to model uncertainty in various domains.

In the field of machine learning, uncertainty is often represented by a probability distribution. The most common approach is to use Bayesian methods, where uncertainty is captured by posterior distributions over model parameters and predictions. Bayesian neural networks, for instance, offer a powerful framework to model uncertainty in deep learning architectures. By incorporating prior beliefs and updating them based on observed data, these networks can produce probabilistic predictions that provide a measure of uncertainty. This is particularly useful in applications where high-confidence predictions are required, and the consequences of errors can be significant.

Recently, another promising approach to uncertainty estimation in machine learning is the use of Monte Carlo Dropout (MC Dropout). MC Dropout leverages the idea of dropout regularization, originally employed during training to prevent overfitting. During prediction, MC Dropout takes multiple forward passes through the network with dropout enabled, which leads to obtaining a distribution of outputs for each input sample. The variance among these sampled predictions serves as a measure of uncertainty. MC Dropout has shown remarkable success in various tasks such as image classification, object detection, and natural language processing. It is computationally efficient and can be easily incorporated into existing deep learning architectures, making it an attractive choice for uncertainty estimation.

Furthermore, there has been a surge of interest in ensemble methods for uncertainty estimation. Ensembles combine the predictions of multiple models to obtain a more robust and calibrated uncertainty measure. Bagging and boosting techniques, which have been widely used in the field of statistics, have found their way into the realm of machine learning for uncertainty estimation as well. By training multiple models with different initializations or using diverse learning algorithms, ensembles can capture different sources of uncertainty, thereby providing a comprehensive assessment of the overall uncertainty in the predictions.

Despite the progress made in uncertainty estimation for machine learning models, challenges still remain. One key concern is the interpretability of uncertainty measures. While probabilistic outputs are more informative than point estimates, effectively communicating uncertainty to end-users and decision-makers is not trivial. Developing visualization techniques and intuitive explanations for uncertainty is an ongoing area of research. Additionally, quantifying uncertainty in high-dimensional data or complex models with massive amounts of parameters requires careful consideration and efficient computational strategies.

In conclusion, uncertainty estimation is a crucial aspect of machine learning models, enabling them to make more informed and reliable predictions. Researchers continue to explore innovative techniques to better represent and quantify uncertainty, allowing for safer and more trustworthy deployment of machine learning systems in real-world scenarios.

Uncertainty is a fundamental part of human experience. Yet, capturing it in a mathematical model is a difficult task. Historically, uncertainty estimation has been a central topic in the field of statistics. Understanding and quantifying uncertainty is crucial in making informed decisions, drawing meaningful conclusions, and assessing the reliability of predictions and inferences derived from data.

In the field of statistics, uncertainty is traditionally addressed through methods like confidence intervals, hypothesis testing, and Bayesian inference. Confidence intervals provide a range of plausible values for a parameter, while hypothesis testing evaluates the significance of differences between groups or variables. Bayesian inference, on the other hand, offers a powerful framework to incorporate prior knowledge and update beliefs based on observed data using probability theory. These approaches have been extensively studied and applied in various domains, from social sciences to natural sciences, to improve the robustness and generalizability of statistical analyses.
However, the advent of machine learning and its widespread adoption in diverse applications has posed new challenges and opportunities in dealing with uncertainty.

Machine learning algorithms are designed to learn patterns from data and make predictions or decisions based on that knowledge. In many cases, it is not enough to provide a single deterministic prediction; it is equally important to convey the level of uncertainty associated with the prediction. For instance, in medical diagnosis, knowing the uncertainty in a model's prediction is crucial when the consequences of a false positive or false negative can be significant. In autonomous driving systems, understanding uncertainty becomes essential to ensure the safety of passengers and pedestrians.

To address the uncertainty challenge in machine learning, researchers have developed various techniques to represent and quantify uncertainty. One of the common approaches is the use of probabilistic models, where uncertainty is captured in the form of probability distributions. Instead of providing a single point prediction, probabilistic models generate a range of possible outcomes along with their associated probabilities. Bayesian neural networks, Gaussian processes, and variational autoencoders are some examples of such models that have gained prominence in the quest for uncertainty-aware machine learning.

Another avenue for dealing with uncertainty in machine learning is through ensemble methods. Ensemble learning combines multiple models to improve predictive accuracy and can also provide uncertainty estimates. By aggregating the outputs of diverse models, ensemble methods can identify situations where predictions are consistent among the models, leading to higher confidence, and cases where the models disagree, indicating higher uncertainty.

As the intersection of statistics and machine learning grows increasingly important, the integration of these two domains has become an active area of research. The incorporation of probabilistic modeling and Bayesian approaches into machine learning methods opens up new possibilities for developing more robust, interpretable, and calibrated models. Moreover, the synergy of statistical techniques with deep learning, reinforcement learning, and other advanced machine learning paradigms has the potential to push the boundaries of uncertainty quantification and application even further.

In this thesis, we aim to explore and contribute to the field of uncertainty in statistics and machine learning. We will investigate state-of-the-art methods for uncertainty estimation, delve into their theoretical underpinnings, and evaluate their performance in various real-world scenarios. By gaining a deeper understanding of uncertainty and its implications, we aspire to pave the way for more reliable, trustworthy, and accountable AI systems in the future.

In the following chapters, we will begin by reviewing the relevant literature on uncertainty in both statistics and machine learning, providing a comprehensive overview of the current landscape. We will then present our proposed methodologies and empirical studies, illustrating how uncertainty quantification can enhance decision-making processes and mitigate risks in practical applications. Finally, we will conclude with a discussion of our findings, potential limitations, and future research directions in this exciting and rapidly evolving field.

\fi
